---
title: "Task 4 - Halloween"
author: "Andrew Wyllie"
output: 
  html_document:
        toc: true
        toc_float: true
---

# Introduction

This data is three consecutive years of results of a Halloween survey conducted by the satirical website boingboing.net. Users were asked to rate how a series of Halloween candies made them feel on a three point scale of "Despair", "Meh" and "Joy. 

The data was provided in three separate Excel spreadsheets containing the data for the years 2015, 2016, and 2017.

# Assumptions made

The nature of the data meant that quite a lot of work was needed to get it in a place for easy analysis, which made a few decisions necessary.

Most columns outside those on the rating scale were open to any input, and as such had some colourful responses.

## Age

Issue: Several people recorded their age as being an exponent (like 1E26) and others made some commentary about being too old or it not being our business.

Decision: Force the column to an integer, coercing some string entries to NA, and also NA any age above 100, because I don't believe for a second anyone a century old is responding to a boingboing Halloween survey. I might have been possible to roughly estimate a few ages if they were vague, but in the interest of time I removed them.

Effect: Likely a slight adjustment to any calculations around respondent age, such as the mean age found in the analysis.

## The "candies"

Issue: Several candies had their names changed over the course of the three years, which may influence the vote by the respondent.

Decision: In many cases I combined these columns for expediency. In one case, "Bonkers" became two separate categories -- Bonkers the candy and Bonkers the board game. I couldn't be totally sure which category to assign the original column to (though I could make an educated guess), so instead opted to keep them separate.

Effect: May lead to some inaccuracies when reporting total favour for specific candies.

Issue: Some of these aren't candy at all!

Decision: I diverged my work. Two separate tables were produced: one in which the entire vote tally is intact, and another that I manually trimmed to only contain what was identifiably what I would consider "candy". This second dataset is what was used to perform analysis.

Result: More useful data. Obviously nobody likes to receive a broken glowstick.

## The miscellaneous questions

Issue: The survey included a grab bag of miscellaneous questions to accompany the main survey. THese are all varying levels of nonsense that the writers wanted to use to make silly statistical comparisons -- people who saw the internet dress as blue and black prefer x candy, for instance.

Decision: Removed from the analysis data, but retained in the main clean data set.

Result: No real effect on the analysis -- it wasn't analysed. Data is intact if any interest arises.

## Countries

Issue: This question field received a bonanza of stupid replies that made analysis difficult.

Decision: Clean the responses as best as possible to allow for better analysis. Much of this is hard-coded.  Because I'm changing answers by my own interpretation, anything that was too ambiguous to feel I could make a call on was filtered as "Other" for the purposes of analysis. Any non-responses were noted as "Not Specified"

Result: Less interesting responses, but hopefully more useful data.

# Cleaning

Oh, boy.

## First steps

Try and unify the columns between the three years. There was a lot of shared names between the first two years, and it was a matter of trying to make the naming common across all three where possible. Some adjustments were made for readability, but I kept the snarkier or more esoteric fields intact where possible.

Step two was to remove the timestamps, but retain the year of each individual dataset to allow comparative analysis later. 2017 has its data formatted somewhat different, so I chose to remove the ID field entirely.

The three tables were then reorganised so that the metadata fields (name, etc.) came first, the candy rating came second, then the miscellany brought up the rear. At this point the three tables were bound together and rearranged one last time.

At this point I wrote a copy of the entire clean dataset to disk, and created a second "simplified" one with the adjustments noted above. No manipulation has been performed on the clean dataset beyond unifying the three years and some column name adjustments.

# Analysis

```{r message=FALSE}
library(tidyverse)
library(here)
```

```{r warning=FALSE, message=FALSE, error=FALSE}
source(here("data_cleaning_scripts/clean_data.R"))
halloween <- read_csv(here("clean_data/halloween_simplified.csv"))
```

### 1. What is the total number of candy ratings given across the three years.

```{r}
halloween %>% 
  summarise(across(7:last_col(), ~sum(!is.na(.)))) %>% 
  mutate(total = rowSums(.)) %>% 
  select(total)
```

Total number of individual candy ratings across the simplified data set is 615,215.

### 2. What was the average age of people who are going out trick or treating?

```{r}
halloween %>% 
  filter(!is.na(age), going_out == "Yes") %>% 
  summarise(avg_age = round(mean(age),1))
```

The average age of people who marked that they intended to go out trick or treating is 35.

### 3. What was the average age of people who are not going trick or treating?

```{r}
halloween %>% 
  filter(!is.na(age), going_out == "No") %>% 
  summarise(avg_age = round(mean(age), 1))
```

The average age of people who marked that they intended to not go out trick or treating is a little over 39.

### 4. For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
halloween %>% 
  select(7:last_col()) %>% 
  pivot_longer(cols = everything(), names_to = "candy", values_to = "rating") %>% 
  group_by(candy, rating) %>% 
  filter(!is.na(rating)) %>% 
  summarise(count = n()) %>% 
  group_by(rating) %>% 
  slice_max(count)
```

For the 3 rating categories, "gum from baseball cards" was rated as "Despair" most often. This is not a surprise as the gum from baseball cards is awful.

"Any full-sized candy bar" received the most joy, and lollipops the most indifference.

### 5. How many people rated Starburst as despair?

```{r}
halloween %>% 
  group_by(starburst) %>% 
  filter(starburst == "DESPAIR") %>% 
  summarise(votes = n())
```

1,990 people rated Starburst as "despair"

### 6. What was the most popular candy bar by this rating system for each gender in the dataset ?

Scoring:

* Joy +1
* Meh 0
* Despair -1

```{r}
halloween %>% 
  select(gender, 7:last_col()) %>% 
  pivot_longer(cols = 2:last_col(), names_to = "candy", values_to = "rating") %>% 
  group_by(candy, rating, gender) %>% 
  filter(!is.na(rating)) %>% 
  mutate(score = recode(rating, "DESPAIR" = -1, "JOY" = 1, "MEH" = 0)) %>% 
  group_by(gender, candy) %>% 
  summarise(score = sum(score)) %>% 
  slice_max(score) %>% 
  arrange(-score)
```

By this scoring method, every category rated "Any full-sized candy bar" as the greatest source of joy. This is not a surprise as who doesn't like a free full-sized candy bar?

### 7. What was the most popular candy bar in each year?

Scoring:

* Joy +1
* Meh 0
* Despair -1

```{r}
halloween %>% 
  select(year, 7:last_col()) %>% 
  pivot_longer(cols = 2:last_col(), names_to = "candy", values_to = "rating") %>% 
  group_by(candy, rating, year) %>% 
  filter(!is.na(rating)) %>% 
  mutate(score = recode(rating, "DESPAIR" = -1, "JOY" = 1, "MEH" = 0)) %>% 
  group_by(year, candy) %>%
  summarise(score = sum(score)) %>% 
  slice_max(score) %>% 
  arrange(year)
```

This result persists when analysing across the three years in the data.

### 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

Scoring:

* Joy +1
* Meh 0
* Despair -1

```{r}
halloween %>% 
  select(country, 7:last_col()) %>% 
  pivot_longer(cols = 2:last_col(), names_to = "candy", values_to = "rating") %>% 
  group_by(candy, rating, country) %>% 
  filter(!is.na(rating)) %>% 
  mutate(score = recode(rating, "DESPAIR" = -1, "JOY" = 1, "MEH" = 0),
         country = case_when(
           str_detect(country, "USA") ~ country,
           str_detect(country, "UK") ~ country,
           str_detect(country, "CANADA") ~ country,
           str_detect(country, "NOT SPECIFIED") ~ country,
           TRUE ~ "OTHER")) %>% 
  group_by(country, candy) %>% 
  summarise(score = sum(score)) %>% 
  slice_max(score) %>% 
  arrange(-score)
```

The result when splitting by country is more interesting. The UK notably favours three candies joint favourite, and none of them are just any full-sized candy bar! Not sure how I feel about our answers, personally.

# Any other insights?

As a quick curiosity, I thought it would be interesting to examine the contents of the additional columns seeking any additional joys and despairs. I imported the full clean data set and exported every phrase (any collection words separated by a comma) from the column.

```{r}
halloween2 <- read_csv(here("clean_data/halloween.csv"))
```

```{r}
halloween2 %>% 
  select(joy_other) %>% 
  mutate(joy_other = case_when(
    str_detect(joy_other, "(?i)bounty") ~ joy_other,
    TRUE ~ NA_character_
  )) %>% 
  filter(!is.na(joy_other))
```

```{r}
joys <- halloween2 %>% 
  select(joy_other) %>% 
  filter(!is.na(joy_other)) %>% 
  pull()
```

### Misc Joy

```{r}
tibble(word(joys, sep = ",")) %>% 
  rename(word = `word(joys, sep = ",")`) %>% 
  mutate(word = str_to_lower(str_replace(word, "\\,", ""))) %>% 
  group_by(word) %>% 
  summarise(count = n()) %>% 
  arrange(-count) %>% 
  slice_max(count, n=10)
```

```{r}
despairs <- halloween2 %>% 
  select(despair_other) %>% 
  filter(!is.na(despair_other)) %>% 
  pull()
```

### Misc despair

```{r}
tibble(word(despairs, sep = ",")) %>% 
  rename(word = `word(despairs, sep = ",")`) %>% 
  mutate(word = str_to_lower(str_replace(word, "\\,", ""))) %>% 
  filter(!word %in% c("anything", "those", "a", "any", "bit", "the")) %>% 
  group_by(word) %>% 
  summarise(count = n()) %>% 
  arrange(-count) %>% 
  slice_max(count, n= 10)
```
It's worth noting that a few of the popular responses ended up as specific fields, so some candies actually have slightly higher Joy ratings than recorded. Almond Joy was such a popular response it actually surprises me it wasn't included in the survey.
